---
title: "Occupational Table Explaratory Data Anylysis"
output: html_notebook
---

# Summary

Stina sent two datasets. One is occupational data for all Canadian immigrants, from 1980 to 2016 (10 and 5 year increments), while the other is a series of archived web-only tables (prepared datasets) for the Haitian Canadian community specifically. I will explore the former here. I don't want to resort to web-scraping the latter. So I will try to gather the Haitian Canadian datasets that underlie the tables in another notebook.

I will on focus on figures that are percentages of overall immigration rates.

```{r setup}
library(tidyverse)
```

## Data import

The csv file for the table that Stina linked had additional information that needed to be remvoed to be parsed correctly:

 "Occupation - National Occupational Classification (NOC) 2016 (693A), Highest Certificate, Diploma or Degree (15), Admission Category and Applicant Type (31), Period of Immigration (7), Age (5A) and Sex (3) for the Immigrant Population Who Landed Between 1980 and 2016 Aged 15 Years and Over, in Private Households of Canada, Provinces and Territories, 2016 Census - 25% Sample Data"

```{r results='hide'}
imm_occ_raw <- read_csv("data/can_imm_mod.csv")
colnames(imm_occ_raw)[1:2] <- c("occupation", "total")
colnames(imm_occ_raw)[3:8] <- str_replace(colnames(imm_occ_raw)[3:8], " to ", "_")
colnames(imm_occ_raw)[8] <- c("2011_2016")

imm_occ_top <- imm_occ_raw[1:3,] %>% 
  add_column(code = NA) %>% 
  relocate(code)

imm_occ_bot <- imm_occ_raw[-c(1:3),] %>% 
  separate(occupation, c("code", "occupation"), sep = "\\s", extra = "merge")

imm_occ <- bind_rows(imm_occ_top, imm_occ_bot) %>% 
  select(-c("X9","2001_2010") )

imm_occ %>% glimpse()
```

## Different views with filtering
```{r}

imm_occ %>%
  filter(str_detect(code, "\\b[1-9]{1}\\b"))

imm_occ %>%
  filter(str_detect(code, "^9[1-9]$"))

imm_occ %>%
  filter(str_detect(code, "^92[1-9]$"))

imm_trans <- imm_occ %>% 
  filter(str_detect(code, "751")) %>% 
  select(matches("\\d"))

```

## Demo Visualization

```{r}
# get transport occupations per year
imm_trans <- imm_occ %>% 
  filter(str_detect(code, "751"))

# plot relative frequencies
bind_cols(imm_trans[c("code", "occupation")], imm_trans %>% 
            transmute(across(matches("\\d"), ~ ./ .[1]))) %>% 
  ggplot()

```

